\documentclass{article}
\usepackage[inline]{asymptote}
\usepackage{enumitem,amsmath,amsfonts,geometry,parskip,amssymb}
\def\asydir{asy}
\title{Division Algorithm, Euclid and Bezout}
\author{Andres Buritica Monroy}
\date{}
\begin{document}
\maketitle
\section{Division Algorithm}
  \begin{itemize}
    \item Prove that if $a$ is an integer and $b$ is a positive integer, there
      is a unique pair $(q,r)$ of integers such that $0\le r<b$ and $a=qb+r$.
    \item Find all integers $n$ such that $2n^2+1\mid n^3+n^2-n-1$.
  \end{itemize}
\section{Euclid's Algorithm}
  We define the \emph{greatest common divisor} of two integers $a$ and $b$, not
  both of which are 0, as the largest positive integer $d$ such that $d\mid a$
  and $d\mid b$. We notate it by $\gcd(a,b)$.
  \begin{itemize}
    \item Let $a$ and $b$ be integers. Prove that if $a=qb+r$ then
      $\gcd(a,b)=\gcd(b,r)$.
    \item Find $\gcd(72,30)$ by applying the previous result repeatedly. (This
      is Euclid's Algorithm.)
    \item Prove that for all positive integers $a,m,n$ we have
      \[\gcd(a^m-1,a^n-1)=a^{\gcd(m,n)}-1.\]
  \end{itemize}
\section{Bezout's Identity}
  \begin{itemize}
    \item Let $a$ and $b$ be integers. Prove that there are integers $c$ and $d$
      such that $ac+bd=\gcd(a,b)$.
    \item Prove that if $n,a,b$ are positive integers such that $n\mid ab$ and
      $\gcd(n,a)=1$ then $n\mid b$.
    \item Let $a,b,c,d$ be positive integers with $\gcd(a,b)=1$. Prove that there is an integer $e$
      such that $a\mid e-c$ and $b\mid e-d$.
  \end{itemize}
\newpage
\section{Problems}
  \begin{enumerate}
    \item The denominators of two irreducible fractions are 600 and 700. Find the
      minimum possible value of the denominator of their sum (written as an irreducible
      fraction).
    \item Let $a,b,c,d$ be positive integers with $ab=cd$. Prove that there
      exist positive integers $p,q,r,s$ such that \(a=pq,b=rs,c=pr,d=qs\).
    \item Prove that for positive integers $m,n>2$ we cannot have \[2^m-1\mid
        2^n+1.\]
    \item Find all positive integers $n$ such that $3^{n-1}+5^{n-1}\mid
      3^n+5^n$.
    \item Let $S$ be a nonempty set of integers such that if $a$ and $b$ are in $S$, then
      so is $2a-b$.

      Prove that $S$ is an arithmetic
      progression. (That is, there are integers $a$ and $d$ such that $n\in S$ if
      and only if 
      $n=a+kd$ for some integer $k$.)
    \item Let $n$ be a composite positive integer. Calculate
      \[\gcd((n-1)!+1,n!).\]
    \item Suppose $a_1,a_2, \dots$ is an infinite strictly increasing sequence of
      positive integers and $p_1, p_2, \dots$ is a sequence of distinct primes
      such that $p_n \mid a_n$ for all $n \ge 1$. It turns out that
      $a_{n+1}-a_n=p_{n+1}-p_n$ for all $n\ge 1$. Prove that the sequence $\{a_n\}$
      consists only of prime numbers.
    \item Assume that $p_1,\ldots,p_m,\ q_1,\ldots,q_n$ are primes such that
      \[p_1p_2\cdots p_m=q_1q_2\cdots q_n.\]
      Prove that the $q_i$s are a permutation of the $p_i$s.
    \item Let $m$ and $n$ be positive integers. Prove that
      \[m\mid\gcd(m,n)\binom mn.\]
  \end{enumerate}
\newpage
\section{Homework}
  \begin{enumerate}
    \item Find all integers $x,y$ such that $6x+4y=8$.
    \item Let $a_1,a_2,\ldots,a_n$ be positive integers, and let $d$ be the
      largest positive integer such that $d\mid a_i$ for all $i$.

      Prove that there are integers $b_1,b_2,\ldots,b_n$ such that
      \[d=a_1b_1+a_2b_2+\cdots+a_n b_n.\]
    \item The Fibonacci sequence is defined by $F_1=F_2=1$ and
      \[F_{n+1}=F_n+F_{n-1}.\]
      Prove that $\gcd(F_n,F_{n+2})=1$ for all $n$.
  \end{enumerate}
\end{document}
