\documentclass{article}
\usepackage[inline]{asymptote}
\usepackage{enumitem,amsmath,amsfonts,geometry,parskip,amssymb}
\def\asydir{asy}
\title{Bounding}
\author{Andres Buritica Monroy}
\date{}
\begin{document}
\maketitle
\section{Techniques}
\begin{itemize}
  \item WLOG arguments: break symmetry to bound one of the variables.
  \item If $a$ and $b$ are integers with $a>b$ then $a\ge b+1$.
  \item If $a$ and $b$ are integers with $b\ne 0$ and $a\mid b$ then
    $|a|\le|b|$. Thus, try to get big dividing small. Polynomial long division
    is helpful here.
  \item Look for ``dominant terms'' that grow larger than everything else. For
    sufficiently large $n$ and any $c>1$, we have $c<\log n<n^c<c^n<n!<n^n$.
  \item If you need to prove a specfic bound of the form ``$f(x)>g(x)$ for
    $x>k$,'' try induction.
  \item If you have a factorial, $n!\equiv 0\pmod c$ for $n\ge c$ so mods are
    often nice.
\end{itemize}
\section{Problems}
\begin{enumerate}
  \item Find all positive integers $k$ such that there is a positive integer $n$
    for which
    \[k=\frac{n^2-29}{3n+11}.\]
  \item Find all positive integers $a,b,c$ such that $ab+bc+ca=abc+2$.
  \item Prove that if $n>11$ then $n^2-19n+89$ is not a square.
  \item Find all positive integers $n$ such that $2^n\le (n+1)^2$.
  \item Find all positive integers $a,b,c$ such that $a\mid b+c,\ b\mid c+a,\
    c\mid a+b$.
  \item Let $S(n)$ be the sum of the digits of $n$. Find $S(S(S(4444^{4444})))$.
  \item  For a natural number $N$, consider all distinct perfect squares that
    can be obtained from $N$ by deleting one digit from its decimal
    representation. Prove that the number of such squares is bounded by some
    value that doesn't depend on $N$.
  \item Find all pairs $a,b$ of positive integers such that $2017^a=b^6-32b+1$.
\end{enumerate}
\newpage
\section{Homework}
\begin{enumerate}
  \item Find all pairs of positive integers $x,y$ such that $x^2-y!=2001$.
  \item Find all pairs of positive integers $x,y$ such that $y^2(x-1)=x^5-1$.
  \item Find all triples $a,b,c$ of positive integers such that $a\mid bc-1,\
    b\mid ca-1$ and $c\mid ab-1$.
\end{enumerate}
\end{document}
