\documentclass{article}
\usepackage[inline]{asymptote}
\usepackage{enumitem,amsmath,amsfonts,amssymb,geometry,parskip}
\def\asydir{asy}
\newcommand\lcm{\text{lcm}}
\title{Prime factorisations and Arithmetic Functions}
\author{Andres Buritica Monroy}
\date{}
\begin{document}
\maketitle
\section{Prime factorisations}
  From last time:
  \begin{itemize}
    \item Assume that $p_1,\ldots,p_m,\ q_1,\ldots,q_n$ are primes such that
      \[p_1p_2\cdots p_m=q_1q_2\cdots q_n.\]
      Prove that the $q_i$s are a permutation of the $p_i$s.
  \end{itemize}
  Therefore, each positive integer has a unique prime factorisation (the
  Fundamental Theorem of Arithmetic).
  In particular we can write a positive integer $n$ uniquely as
  \[n=p_1^{e_1}p_2^{e_2}\cdots p_k^{e_k},\]
  where $p_i$ are all prime and $e_i$ are all positive integers.

  Prime factorisations allow us to view statements about divisibility and
  multiplication in terms of the exponents $e_i$. Remember that when you
  multiply two numbers together, you add the exponents on the corresponding
  primes. For example,
  \[2^2 3^1 5^1\times 2^3 7^2=2^5 3^1 5^1 7^2.\]

  In what follows, let 
  \[a=p_1^{e_1}p_2^{e_2}\cdots p_m^{e_m},\
        b=q_1^{f_1}q_2^{f_2}\cdots q_k^{e_k}.\]
  \begin{itemize}
    \item $a\mid b$ if and only if for each $i$ we have that $p_i=q_j$
      for some $j$, and that $e_i\le f_j$.
    \item $a$ is a perfect $k$th power if and only if $k\mid e_i$ for all $i$.
    \item The lcm is found by taking the maximum power of each prime that
      divides either $a$ or $b$; the gcd is found by taking the minimum power of
      each prime that divides both $a$ and $b$.
    \item $\gcd(a,b)\times\lcm(a,b)=ab$.
  \end{itemize}
\section{Arithmetic functions}
  We define:
  \begin{itemize}
    \item The number of positive divisors function $d(n)$.
    \item The sum of positive divisors function $\sigma(n)$.
    \item The totient function $\varphi(n)$: the number of positive integers
      which are at most $n$ and coprime to $n$.
  \end{itemize}
  Find formulae for $d(n)$ and $\sigma(n)$ in terms of the prime factorisation
  of $n$.
\section{Problems}
\begin{enumerate}
  \item We call the \emph{main divisors} of a composite number $n$ the two
    largest of its divisors other than $n$. Composite numbers $a$ and $b$ are
    such that their main divisors coincide. Prove that $a=b$.
  \item Prove that $d(n)\le 2\sqrt n$.
  \item Let $a,b,p$ be positive integers such that $p$ is prime and
    $\lcm(a,a+p)=\lcm(b,b+p)$. Prove that $a=b$.
  \item Prove that for all $n$,
    \[\sigma(1)+\sigma(2)+\cdots+\sigma(n)\le n^2.\]
  \item Prove that if $ab$ is a perfect square, then so are $\frac a{\gcd(a,b)}$ and
      $\frac b{\gcd(a,b)}$.
  \item Prove that for all composite $n$ apart from 6,
    \[\sqrt n\le\varphi(n)\le n-\sqrt n.\]
  \item Determine all composite integers $n>1$ that satisfy the following
    property: if $d_1, d_2, \ldots, d_k$ are all the positive divisors of $n$
    with $1=d_1<d_2<\cdots<d_k=n$, then $d_i$ divides $d_{i+1}+d_{i+2}$ for
    every $1 \leq i \leq k-2$.
  \item Let $n$ be an even positive integer such that $\sigma(n)=2n$. Prove that
    $n=2^{p-1}\left(2^p-1\right)$, where $p$ is a prime.
\end{enumerate}
\newpage
\section{Homework}
\begin{enumerate}
  \item The cells in a jail are numbered from 1 to 100, and there are 100
    buttons also numbered from 1 to 100. For each $i$, the $i$th button opens a
    closed cell and closes an open cell, affecting only the multiples of $i$. 

    For example, if the 47th cell is open and the 94th cell is closed, then
    pressing the 47th button will close the 47th cell and open the 94th cell.

    Initially all cells are closed.
    The warden presses the first button, then the second, and so on, for all
    100 buttons. Which cells are open at the end?
  \item Find all primes $p$ such that $p^{2022}+p^{2023}$ is a perfect square.
  \item Prove that for any positive integer $n$ we have $\sigma(n)\ge d(n)\sqrt
    n$.
\end{enumerate}
\end{document}
