\documentclass{article}
\usepackage[inline]{asymptote}
\usepackage{enumitem,amsmath,amsfonts,geometry,parskip,amssymb}
\def\asydir{asy}
\newcommand\lcm{\mathrm{lcm}}
\newcommand\Zz{\mathbb{Z}}
\newcommand\ord{\mathrm{ord}}
\title{Review and Extension: Primes}
\author{Andres Buritica}
\date{}
\begin{document}
\maketitle
\section{Exam review}
\newpage
\section{Quadratic residues}
Assume there exists a generator mod $n$.
An element $x\in\Zz_n^*$ can be written as $y^k$ for $y\in\Zz_n^*$ iff
$\ord_n(x)\gcd(\varphi(n),k)\mid \varphi(n)$.

If $x$ can be written as $y^2$ for $y\in\Zz_n^*$, then we say that $x$ is a
\emph{quadratic residue (QR)} mod $n$. Note that $0$ is not a QR mod $n$.

Prove that $x$ is a QR mod $n$ iff both
\begin{itemize}
  \item $x\equiv 1\pmod{\gcd(8,n)}$, and
  \item for each odd prime $p\mid n$, $x$ is a QR mod $p$.
\end{itemize}

Hence, we now restrict ourselves to considering QRs mod $p$.
Define the \emph{Jacobi symbol}
\[\left(\frac ap\right)=\begin{cases} 0&p\mid a \\ 1&a\text{ is a QR mod }p \\
-1&\text{otherwise}\end{cases}.\]

\begin{itemize}
  \item (Euler's criterion) Prove that
    $\left(\frac ap\right)\equiv a^{\frac{p-1}2}\pmod p$.
    Hence, $\left(\frac ap\right)\left(\frac
    bp\right)=\left(\frac{ab}p\right)$.
  \item (Gauss' Lemma) Let $a\in\Zz_p^*$, and let $S\subseteq\Zz_p^*$ such that $x\in S\iff
    -x\not\in S$. Let $T=\{ay:y\in S\}$. Then
    $\left(\frac ap\right)=(-1)^{|T\setminus S|}$.
  \item Find $\left(\frac 2p\right)$ and $\left(\frac{-1}p\right)$.
  \item Find all completely multiplicative functions $f:\mathbb N\to\mathbb R$
    such that for all $a,b\in\mathbb N$, at least two of $f(a),f(b),f(a+b)$ are
    equal.
  \item (Quadratic Reciprocity) Let $p$ and $q$ be distinct odd primes. Using
    Gauss' Lemma, prove that
  \[\left(\frac qp\right)\left(\frac pq\right)=(-1)^{\frac{(p-1)(q-1)}4}.\]
\end{itemize}
\section{Problems}
\begin{enumerate}
  \item Let $p$ be an odd prime. Prove that
    \[p^2\mid 1^p+2^p+\cdots+p^p.\]
  \item Let $a,b,c$ be positive integers such that $c\mid a^c-b^c$. Prove that
    $c(a-b)\mid a^c-b^c$.
  \item Prove that for all positive integers $n$,
    \[\binom{2n}n\mid\lcm(1,2,\ldots,2n).\]
  \item Find all pairs of positive integers $x,p$ such that $p$ is prime, $x\le
    2p$, and $x^{p-1}\mid (p-1)^x+1$.
  \item Let $a,b,n$ be positive integers such that $a>b>1$ and $b$ is odd. If
    $b^n\mid a^n-1$, prove that $na^b>3^n$.
  \item Find all natural numbers $n$ such that $n^2\mid 2^n+1$.
  \item Find all positive integers $n$ such that $n\mid 2^n-1$.
  \item Prove that if $\sigma(n)=2n+1$, then $n$ is a perfect square.
  \item Find all positive integers $n$ such that $n\mid 2^{n-1}+1$.
  \item Find all primes $p,q,r$ such that $p\mid q^r+1,\ q\mid r^p+1,\ r\mid
    p^q+1$.
  \item Let $n$ and $m$ be nonnegative integers, and let $p$ be prime. Prove
    that
    \[\binom nm\equiv\prod_{i=0}^k\binom{n_i}{m_i}\pmod p,\]
    where $n=\sum n_i p^i$ and $m=\sum m_i p^i$.
  \item Find all positive integers $n$ for which there exists a function
    $g:\Zz_n\to\Zz_n$ such that all the functions
    \[g(x),g(x)+x,\ldots,g(x)+100x\]
    are bijections $\Zz_n\to\Zz_n$.
  \item Let $k$ and $n$ be positive integers such that $n$ is odd. Prove that
    there is an integer $a$ such that $a^{32}\equiv(n+1)^3\pmod{n^k}$.
  \item Prove that for any prime $p$ and positive integer $a$ with $p\nmid a$
    there are at least $p-1$ solutions in $\Zz_p$ to $x^2+y^2\equiv a\pmod p$.
  \item Let $p$ be prime and let $a$ and $b$ be positive integers such that
    $p\nmid b$. Prove that there exists a positive integer $n$ such that
    $p^a\mid n^n-b$.
  \item If $p>3$ is a prime such that $\varphi(p-1)>\frac{p-1}3$, prove that
    there are two consecutive generators mod $p$.
  \item Let $p$ be a prime and let $n$ be a positive integer with $p\nmid n$.
    Find \[\sum_{i=1}^p\left(\frac{i^2+n}p\right).\]
  \item Let $p>3$ be a prime and let $a,b,c$ be integers with $a\ne 0$. Suppose
    that $ax^2+bx+c$ is a perfect square for $p$ consecutive integers $x$. Prove
    that $p\mid b^2-4ac$.
  \item Let $P$ be a nonconstant polynomial with integer coefficients. Prove that for any
    integer $m$ there exist an integer $n$ and a prime $p$ such that $p^m\mid
    P(n)$.
  \item Let $a_1,a_2,a_3,\ldots$ be a sequence of integers, such that for any
    positive integers $n$ and $k$, the quantity
    \[\frac{a_n+a_{n+1}+\cdots+a_{n+k-1}}k\]
    is always the square of an integer. Prove that all $a_i$s are equal.
\end{enumerate}
\section{Homework}
Solve and submit any 3 problems from section 3.
\end{document}
