\documentclass{article}
\usepackage[inline]{asymptote}
\usepackage{enumitem,amsmath,amsfonts,amssymb,geometry,parskip,rotating}
\def\asydir{asy}
\title{Review and Extension --- Size Arguments}
\author{Andres Buritica}
\date{}
\begin{document}
\maketitle
\section{Exam review}
\section{Key concepts for this term}
  \begin{itemize}
    \item WLOG
    \item Big divides small
    \item Dominant terms
    \item Graphical reasoning
    \item Relative growth rates of common functions
    \item Infinite descent
    \item Vieta jumping
    \item Pell equations
    \item Counting numbers in a range
    \item Dealing with floor functions
  \end{itemize}
\section{Problems}
\begin{enumerate}
  \item Find all pairs of positive integers $x,y$ such that
    $1+2^x+2^{2x+1}=y^2$.
  \item Find all pairs of positive integers $x,y$ such that $x^3-y^3=xy+61$.
  \item Find all triples of positive integers $a,b,c$ such that $a^2+b+c=abc$.
  \item Find all positive integers $x,y,n$ such that $\gcd(x,n+1)=1$ and
    $x^n+1=y^{n+1}$.
  \item Four positive integers $x,y,z,t$ satisfy $xy-zt=x+y=z+t$. Is it possible
    that $xy$ and $zt$ are both perfect squares?
  \item Prove that there are infinitely many triples $(a,b,c)$ of positive
    integers in arithmetic progression
    such that $ab+1$, $bc+1$ and $ca+1$ are all perfect squares.
  \item Find all solutions in integers to $x^2+y^2+z^2=2xyz$.
  \item Let $a$ and $b$ be two positive integers. Prove that
    \[a^2+\left\lceil\frac{4a^2}b\right\rceil\] is not a square.
  \item Let $p$ and $q$ be coprime. Prove that
    \[\sum_{i=1}^{q-1}\left\lfloor\frac{ip}{q}\right\rfloor=\frac{(p-1)(q-1)}2.\]
  \item Find all positive integers $n$ such that $1+\lfloor\sqrt n\rfloor$
    divides $n$.
  \item Let $r$ be an irrational root of a polynomial $P(x)$ of degree $d$ with
    integer coefficients. Prove that there is a real number $C$ such that for
    any integer $q$ we have $\{qr\}\ge \frac C{q^{d-1}}$.
  \item Let $x$ be an irrational number.
    \begin{enumerate}
      \item Prove that for each positive integer $n$ there is a positive integer
        $m$ such that $\{mx\}<\frac 1n$.
      \item Prove that there are infinitely many positive integers $n$ such
        that $\{nx\}<\frac 1n$.
    \end{enumerate}
  \item Prove that the sequence $a_i=\left\lfloor(\sqrt 2+1)^i\right\rfloor$
    alternates between even and odd integers.
  \item Find all positive integers $x$ and $y$ such that if $z=\gcd(x,y)$, then
    $x+y^2+z^3=xyz$.
  \item For each positive integer $n$, let
    \[f(n)=\frac1n\sum_{i=1}^n \left\lfloor\frac ni\right\rfloor.\]
    \begin{enumerate}
      \item Prove that $f(n+1)>f(n)$ for infinitely many $n$.
      \item Prove that $f(n+1)<f(n)$ for infinitely many $n$.
    \end{enumerate}
  \item Is there a positive integer $m$ for which the equation
    \[\frac1a+\frac1b+\frac1c+\frac1{abc}=\frac m{a+b+c}\]
    has infinitely many solutions in positive integers $a,b,c$?
\end{enumerate}
\section{Homework}
Solve and submit any three problems from Section 3.
\newpage
\section{Extension: density of the primes}
\begin{enumerate}
  \item
    Let $n$ be a positive integer larger than 1.
    \begin{enumerate}
      \item Prove that the product of all primes between $\left\lceil\frac
        n2\right\rceil$ and $n$
        (including $n$, not including $\left\lceil\frac n2\right\rceil$)
        is less than
        $2^n$.\footnote{\begin{turn}{180}Hint: binomial coefficients are
        integers.\end{turn}}
      \item Prove that the product of all primes between 1 and $n$ is at most
        $4^{n-1}$.
      \item Find some real number $c$ independent of $n$ such that there are at
        most $\frac{cn}{\log_2 n}$ primes that are at most $n$.
    \end{enumerate}
  \item 
    Let $n$ be a positive integer larger than $2^{2^{2^2}}$.
    \begin{enumerate}
      \item Let $p$ be a prime.
        \begin{itemize}
          \item Prove that if $p^k\mid\binom{2n}n$ then $p^k<
            2n$.
          \item Prove that if $2p\le 2n< 3p$ then $p\nmid\binom{2n}n$.
        \end{itemize}
      \item Prove that 
        \[\prod_{\substack{p^k\|\binom{2n}n\\ p\le n}}p^k<\binom{2n}n.\]
      \item Find some real number $c$ independent of $n$ such that there are at
        least $\frac{cn}{\log_2 n}$ primes that are at most $n$.
    \end{enumerate} 
\end{enumerate}
\end{document}
