\documentclass{article}
\usepackage[inline]{asymptote}
\usepackage{enumitem,amsmath,amsfonts,geometry,parskip,amssymb}
\def\asydir{asy}
\title{Induction and Divisibility}
\author{Andres Buritica Monroy}
\date{}
\begin{document}
\maketitle
\section{Induction and variants}
Arguably, the defining property of the integers is the \textbf{Principle of
	Mathematical Induction}: if we have a set $S\subseteq\mathbb N$ such that
$1\in S$ and $\forall a\in S,\ a+1\in S$ then $S=\mathbb N$.

To prove that some sentence $P(n)$ is true for all positive integers $n$, we
follow the following structure.
\begin{itemize}
	\item Prove that $P(1)$ is true.
	\item Prove that if $P(n)$ is true, then $P(n+1)$ is true.
\end{itemize}
If we've done both of those things, why can we conclude that $P(a)$ is true
for all $a\in\mathbb N$?

We may use induction to prove some foundational results about the integers:
\begin{itemize}
	\item For all positive integers $n$, $n\ge 1$.
	\item If $S$ is a nonempty set of positive integers, there is some $a\in
		      S$ such that for any $b\in S$ we have $a\le b$.
	      This is known as the \textbf{Well-Ordering Principle}.
\end{itemize}

The final addition to our induction toolkit will be \textbf{strong induction}:
if $S$ is a set of positive integers such that $1\in S$ and
\[\forall a\in \mathbb N,\ (\forall b\in\mathbb N,\ b\le a\implies b\in
	S)\implies a+1\in S,\]
then $\mathbb N=S$.
\section{Divisibility}
For integers $a$ and $b$, we say $a\mid b$ (read ``$a$ divides
$b$'') if there is some integer $c$ with $b=a\times c$.
\begin{itemize}
	\item Prove that if $a\mid b$ and $a\mid c$, then for any integers $m$ and
	      $n$, $a\mid bm+cn$.
	\item Prove that if $a$ and $b$ are positive integers with $a\mid b$ then
	      $a\le b$.
\end{itemize}

We define a \emph{prime} as a positive integer larger than 1 which is not
divisible by any positive integer other than 1 and itself.
\begin{itemize}
	\item Prove that every positive integer larger than 1 has a prime factor.
	\item Prove that there are infinitely many primes.
\end{itemize}
\section{GCD, Euclid, Bezout}
We define the \emph{greatest common divisor} of two integers $a$ and $b$, not
both of which are 0, as the largest positive integer $d$ such that $d\mid a$
and $d\mid b$. We notate it by $\gcd(a,b)$. For convenience we also define
$\gcd(0,0)=0$.

Similarly, the \emph{least common multiple} $\mathrm{lcm}(a,b)$ is the
smallest positive integer $l$ such that $a\mid l$ and $b\mid l$. For
convenience we also define $\gcd(0,a)=0$.
\begin{itemize}
	\item Let $a$ be an integer and let $b$ be a positive integer. Prove that
	      there is exactly one pair of integers $(q,r)$ with $0\le r<b$ such that
	      $a=qb+r$. (Division Algorithm)
	\item Prove that for any integers $a,b,q,r$, if $a=qb+r$ then
	      $\gcd(a,b)=\gcd(b,r)$. (Euclid's Algorithm)
	\item Let $a$ and $b$ be integers. Prove that there are integers $c$ and $d$
	      such that $ac+bd=\gcd(a,b)$. (Bezout's Identity)
	\item Let $a$ and $b$ be integers with $\gcd(a,b)=1$, and let $c$ be an
	      integer. Prove that if $a\mid c$ and $b\mid c$ then $ab\mid c$.
	\item Assume that $p_1,\ldots,p_m,\ q_1,\ldots,q_n$ are primes such that
	      \[p_1p_2\cdots p_m=q_1q_2\cdots q_n.\]
	      Prove that the $q_i$s are a permutation of the $p_i$s.
\end{itemize}
\section{Prime Factorisations}
Therefore, each positive integer has a unique prime factorisation (the
Fundamental Theorem of Arithmetic).
In particular we can write a positive integer $n$ uniquely as
\[n=p_1^{e_1}p_2^{e_2}\cdots p_k^{e_k},\]
where $p_i$ are all prime and $e_i$ are all positive integers.

Prime factorisations allow us to view statements about divisibility and
multiplication in terms of the exponents $e_i$.

In what follows, let
\[a=p_1^{e_1}p_2^{e_2}\cdots p_m^{e_m},\
	b=q_1^{f_1}q_2^{f_2}\cdots q_k^{e_k}.\]
\begin{itemize}
	\item Prove that $a\mid b$ if and only if for each $i$ we have that $p_i=q_j$
	      for some $j$, and that $e_i\le f_j$.
	\item Prove that $a$ is a perfect $k$th power if and only if $k\mid e_i$ for all $i$.
	\item Prove that the lcm is found by taking the maximum power of each prime that
	      divides either $a$ or $b$, while the gcd is found by taking the minimum power of
	      each prime that divides both $a$ and $b$.
	\item Prove that $\gcd(a,b)\times\mathrm{lcm}(a,b)=ab$.
\end{itemize}
\section{Strategies}
\begin{itemize}
	\item Take out the gcd: that is, if you have two integers $a$ and $b$ you
	      can write $a=dx,\ b=dy$ where $x$ and $y$ are coprime.
	\item Factorisations: two especially useful ones are
	      \begin{align*}
		      axy+bx+cy=d\iff (ax+c)(ay+b)=ad+bc, \\
		      a^k-b^k=(a-b)\left(a^{k-1}+a^{k-2}b+\cdots+b^{k-1}\right).
	      \end{align*}
\end{itemize}
\section{Problems}
\begin{enumerate}
	\item Find all integers $n$ such that $n^2+1$ divides $n^3+n^2-n-15$.
	\item Given three distinct natural numbers $a,b,c$, show that
	      \[\gcd(ab+1,bc+1,ca+1)\le\frac{a+b+c}3.\]
	\item Let $p$ be a prime with $p>3$. Prove that there are positive integers
	      $a<b<\sqrt p$ such that $p-b^2\mid p-a^2$.
	\item Let $S$ be the set of ordered pairs of integers. We say that two
	      elements $(a,b)$ and $(c,d)$ of $S$ are $k$-friends if there is an element
	      $(e,f)$ of $S$ such that the area of the triangle formed by these three
	      points is $k$.\footnote{The \emph{shoelace formula} states that the area
		      of this triangle is $\frac12(ad-bc+cf-de+eb-af)$.}
	      Find the smallest positive integer $k$ such that there exists a set of
	      $200$ elements of $S$ such that any pair of them are $k$-friends.
	\item Find all pairs of positive integers $a,b$ such that \[b^2-a\mid
		      a^2+b\quad\text{and}\quad a^2-b\mid b^2+a.\]
	\item Prove that for any nonnegative integer $n$, the number $7^{7^n}+1$ is
	      the product of at least $2n+3$ (not necessarily distinct) primes.
\end{enumerate}
\newpage
\section{Homework}
\begin{enumerate}
	\item Prove that every positive integer can be uniquely represented as a sum
	      of one or more Fibonacci numbers such that the sum does not include two
	      consecutive Fibonacci numbers.
	\item Let $a, b, c$ be positive integers with $a^2+b^2=c^2$, such that no
	      positive integer larger than 1 divides all of them. Prove that
	      there exist positive integers $x,y,z$ such that $a,b,c$ equal
	      $x^2-y^2,2xy,x^2+y^2$ in some order.
	\item
	      \begin{enumerate}
		      \item Find all integers $a,b,c$ with $1<a<b<c$ such that $(a-1)(b-1)(c-1)$
		            divides $abc-1$.
		      \item Do there exist distinct prime numbers $a,b,c$ such that
		            \[a\mid bc+b+c,\ b\mid ac+a+c,\ c\mid ab+a+b?\]
	      \end{enumerate}
\end{enumerate}
\end{document}
